---
title: "PA1_template"
author: "Christopher Kuo"
date: "May 9, 2017"
output: md_document
---

1. Code for reading in the dataset and/or processing the data

```{r, message = FALSE, warning = FALSE}
require(dplyr)
t <- read.csv("activity.csv")
tdate <- group_by(t,date)
tdatesum <- summarize(tdate, sum(steps, na.rm = TRUE))
tdatesum$date <- as.Date(tdatesum$date)
```

2. Histogram of the total number of steps taken each day

```{r plot1, warning = FALSE, message = FALSE}
require(ggplot2)

p <- ggplot(tdatesum, aes(x = date, y=`sum(steps, na.rm = TRUE)`)) +geom_bar(stat = "identity") + scale_x_date()
p2 <- p + ylab("Total Steps") + xlab("Date") + ggtitle("Total Steps Per Day")
p2
```

3. Mean and median number of steps taken each day
```{r}
meansteps <- mean(tdatesum$`sum(steps, na.rm = TRUE)`, na.rm= TRUE)
meansteps
mediansteps <- median(tdatesum$`sum(steps, na.rm = TRUE)`)
mediansteps
```

4.Time series plot of the average number of steps taken
```{r plot2}
tinterval <- group_by(t, interval)
tintervalmean <- summarize(tinterval, mean(steps, na.rm = TRUE))
pint = ggplot(tintervalmean, aes(x=interval, y= `mean(steps, na.rm = TRUE)`)) + geom_line()
pint2 <- pint + ylab("Average Steps") + xlab("Interval Number") + ggtitle("Average Steps Taken Per Time Interval")
pint2
```

5.The 5-minute interval that, on average, contains the maximum number of steps
```{r}
which.max(tintervalmean$`mean(steps, na.rm = TRUE)`)
```

6. Code to describe and show a strategy for imputing missing data

Fill in missing values with average value at that interval across all days.
```{r}
table(is.na(t$steps))


timputed <- t
impute <- function(steps, interval){
    corsteps <- NA
    if(is.na(steps) == TRUE){
        corsteps <- tintervalmean$`mean(steps, na.rm = TRUE)`[which(tintervalmean$interval == interval)]
    }
    else{
        corsteps <- steps
    }
    corsteps
}
timputed$steps <- mapply(impute, timputed$steps, timputed$interval)
table(is.na(timputed$steps))
```

7.Histogram of the total number of steps taken each day after missing values are imputed
```{r plot3}
tdate <- group_by(timputed,date)
tdatesum <- summarize(tdate, sum(steps, na.rm = TRUE))
tdatesum$date <- as.Date(tdatesum$date)

p <- ggplot(tdatesum, aes(x = date, y=`sum(steps, na.rm = TRUE)`)) +geom_bar(stat = "identity") + scale_x_date()
p2 <- p + ylab("Total Steps") + xlab("Date") + ggtitle("Total Steps Per Day With Imputation")
p2
```

8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r plot4, message = FALSE, warning = FALSE}
library(lubridate)
t$date <- as.Date(t$date)
t$day <- wday(t$date)

is.weekday <- function(day){
    if(day == 1 | day == 7){
        weekday <- "weekend"
    }
    else{
        weekday <- "weekday"
    }
    weekday
}

t$weekday <- t$day
t$weekday <- sapply(t$day, is.weekday)

tintday <- group_by(t, interval, weekday)
tintdaymean <- summarize(tintday, mean(steps, na.rm = TRUE))
pintday = ggplot(tintdaymean, aes(x=interval, y= `mean(steps, na.rm = TRUE)`)) + geom_line()
pintday2 <- pintday + ylab("Average Steps") + xlab("Interval Number") + ggtitle("Average Steps Taken Per Time Interval Divided by Weekday")
pintday3 <- pintday2 + facet_grid(weekday ~ .)
pintday3
```